### 2018-02-27 figs/motivation/content.tex

\begin{figure}[b!]
\centering
\includegraphics[height=14em]{figs/motivation/content}
\caption{
How the projects of my dissertation fit into the categorization of explanations
as shown in \figref{figs:motivation_expl_classes}.
As input data is focused on data pre-processing it doesn't appear in the
original figure. Furthermore, no works focusing on structure explanations are
presented as they require knowledge about the internal state of the model.
However, with \textbf{Class Signatures} model induction based on the original
model's outputs and subsequent structure explanations of the induced model are
part of the work-flow.
}
\label{figs:motivation_content}
\end{figure}