### 2018-02-27 figs/new_parallel.tex

\begin{figure}[t]
\centering
\includegraphics[height=16em]{figs/new_parallel.png}
\caption{
The extended interface to allow for non-binary data.
The list on the left shows the discovered ``decision sets".
Below that the distribution of prediction scores are shown.
A parallel coordinate view shows the instances covered by such a decision set (``expl 8") allowing non-binary input data.
The user highlights ``expl 10" showing its instances in purple.
}
\vspace*{-1em}
\label{figs:new_parallel}
\end{figure}