### 2018-02-27 figs/seekaview/workflow.tex

\begin{figure}[t]
\centering
\includegraphics[width=0.8\columnwidth]{figs/seekaview/workflow}
\caption{Framework for guided reconfiguration of interesting views out of high-dimensional data: A tight integration of system generated suggestions with analysts' actions leads to flexibility in having multiple starting points of analyses~(A and B), and transparency in being able to refine the suggestions and adapt both dimension and item space interactively.
}
\label{figs:seekaview_workflow}
\end{figure}