### 2018-02-27 prospector/ui.tex

\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{prospector/ui_inspect} % 0.775
\\
\vspace*{0.75em} % 2
\includegraphics[width=0.3\linewidth]{prospector/ui_sugg} % 0.3
\hspace*{0.2\linewidth} % 18
\includegraphics[width=0.3\linewidth]{prospector/ui_change} % 0.3
% \vspace*{1em}
\caption{
The user interface of \prospector is shown at the top.
The bottom left shows suggestions on what changes (white circles) would decrease the
predicted risk the most.
The bottom right shows how the color plots change due to changing a value (namely changing the
bmi value from 0 to 1).
Fully white circles show the original value of the given patient.
}
\label{figs:ui}
\end{figure}