### 2018-02-27 tex/introduction/usecase.tex

\section{Use Case: Machine Learning for Health Care}
\todo{health}
* inconsistent coding NDC, IDC, hierarchies
* Doctor decision tree