### 2018-02-27 figs/motivation/expl_classes.tex

\begin{figure}
\centering
\includegraphics[height=11em]{figs/motivation/expl_classes}
\caption{
The relation of output-, interaction-, and structure-explanations.
With the help of model induction over a reduced or carefully selected
input, output- and interaction-explanations can utilize structure explanations
of simple machine learning models.
Output- and interaction-explanations can be created while treating the model
as black-box. This is not possible for structure explanations.
}
\label{figs:motivation_expl_classes}
\end{figure}