### 2018-02-27 figs/prospector/risk.tex

\begin{figure}
\centering
\begin{minipage}[c]{0.35\textwidth}
\includegraphics[width=\linewidth]{figs/prospector/dec_risk_compact} % 19.5
\end{minipage}\hfill
\begin{minipage}[c]{0.6\textwidth}
\caption{
Identifying changes to features that reduce the risk of a high risk patient.
The features are sorted by decreasing impact to make large sets of features manageable.
The background color of each feature indicates the predicted risk for this value.
Impossible values, if known, appear grayed out and the slider snaps back to the last possible value if selected.
}
\label{figs:prospector_risk}
\end{minipage}
\end{figure}