### 2018-02-27 old.tex

% \documentclass[60pt]{ociamthesis}  % default square logo 
% \documentclass[12pt,beltcrest]{ociamthesis} % use old belt crest logo
\documentclass[12pt]{ociamthesis} % use older shield crest logo

%load any additional packages

%input macros (i.e. write your own macros file called mymacros.tex 
%and uncomment the next line)
%\include{mymacros}

\title{Using Visual Analytics to Explain\\[1ex]
       Black-Box Machine Learning}   %note \\[1ex] is a line break in the title

\author{\textbf{Josua Walter Hugo Krause}}             %your name
\college{Tandon School of Engineering}  %your college

%\renewcommand{\submittedtext}{change the default text here if needed}
\degree{Doctor of Philosophy, Computer Science}     %the degree
\degreedate{\today}%{November 23, 2016}         %the degree date

%end the preamble and start the document
\begin{document}

%this baselineskip gives sufficient line spacing for an examiner to easily
%markup the thesis with comments
\baselineskip=18pt plus1pt

%set the number of sectioning levels that get number and appear in the contents
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

\maketitle                  % create a title page from the preamble info
% \include{dedication}        % include a dedication.tex file
% \include{acknowlegements}   % include an acknowledgements.tex file
% \include{abstract}          % include the abstract

\input{tex/abstract}
\begin{romanpages}          % start roman page numbering

\tableofcontents            % generate and include a table of contents
% \listoffigures              % generate and include a list of figures
\end{romanpages}            % end roman page numbering

%now include the files of latex for each of the chapters etc
\input{tex/introduction}
\input{tex/infuse}
\input{tex/prospector}
\input{tex/explainer}
\input{tex/current}

%now enable appendix numbering format and include any appendices
% \appendix
% \include{appendix1}
% \include{appendix2}

%next line adds the Bibliography to the contents page
\addcontentsline{toc}{chapter}{Bibliography}
%uncomment next line to change bibliography name to references
%\renewcommand{\bibname}{References}
\bibliography{main}        %use a bibtex bibliography file refs.bib
\bibliographystyle{plain}  %use the plain bibliography style

\end{document}