### 2018-02-27 tex/introduction/motivation.tex

\section{Motivation}
\todo{Motivation}
Modern machine learning models

* 3 failure modes: hidden bias, missing context, incomplete generalization
* failure modes different than for humans / surprising to humans
* 3 examples: pneumonia, radio, toaster
% \cite{Caruana:2015:IMH:2783258.2788613}

% \cite{Bird:2002:ERI:1251972.1252349}
% https://people.duke.edu/~ng46/topics/evolved-radio.pdf

% \cite{2017arXiv171209665B}
% https://arxiv.org/pdf/1712.09665.pdf

* reasons for black box
* gradient to explanation