### 2018-04-02 aggexplain/discussion.tex

\section{Discussion}
\label{sec:ag_discussion}
We showed that aggregating instance-level explanations can be an effective way of enabling humans to identify biases in the input data of machine learning tasks.
Even though, assisted aggregation is as effective as unassisted individual inspection it scales better with large data sets.
Individually inspecting instances in the data is only possible on a sample of the data and requires extrapolation of findings to the whole data set.
Aggregation does not suffer from this, as the representation of the data is independent from its size.
Even though, it requires dedication, our test data set was small enough to still be able to scan in full if necessary.

Furthermore, the bias planted in the data was simple enough to be able to be found in \emph{all} conditions.
This might not be true for real-world data sets with more complex biases.
Even though, histograms are advantageous with respect to tables in finding arbitrary patterns, they are still limited to only one dimension.
Biases that are present only through combinations of features will not be detectable.

In our study, we could confirm findings from Stumpf~\etal~\cite{harmful} and overcome their limitations by using instance-level explanations with aggregation.
However, we could not overcome trust in machine learning model authority, despite being confronted with contradictory evidence, in all cases.
Speculatively, this might stem from people being used to being presented with cleaned up and validated data, as this cumbersome process is often hidden from the end result.

% On this note, in order to guarantee that our data is of value, we experienced a high churn rate of eligible participants ($\geq1.16$ per eligible participant), even though the task seemed relatively simple, from the perspective of visual analytics researchers. \todo{meh, sounds braggy...} \todo{Do we even need to mention this?  I'd just remove.}

\section{Conclusion \& Future Work}
\label{sec:conclusion}
We presented a novel way of aggregating and comparing instance-level explanations.
We found that this method can help humans identify biases in the input data to machine learning models.
However, this is only the case in combination.
Aggregation alone or individual instance-level explanations might lead to worse performance in this regard.
We demonstrated, that an aggregated instance-level explanation approach is as effective as going through the data unassisted.
This is promising, as the proposed method is independent of the size of the data set and thus likely more scalable than its non-aggregated counterpart.
However, specifically confirming this hypothesis remains future work.

As we were conducting an exploratory analysis of the study, individual findings remain to be tested in-situ in future work.
Furthermore, experimenting with more complex forms of data biases opens up additional research opportunities.

All in all, we presented a usable method for effectively utilizing instance-level explanations on a large scale.
As machine learning models become more complex and opaque, this becomes an important stepping stone in tackling the behemoth of effectively improving machine learning models and their data alike.