### 2018-04-02 tex/aggexplain.tex

\chapter{User Study on Aggregating Instance-level Explanations}
\label{chap:aggexplain}
\input{aggexplain/template.tex}

\section{General Discussion}
\todo{TODO}