### 2018-04-03 prospector/compress.tex

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{prospector/compress_1} \\  \vspace*{0.2em} % 0.8 no vs
\includegraphics[width=0.9\linewidth]{prospector/compress_2} \\  \vspace*{0.5em} % 0.9 no vs
\includegraphics[width=0.7\linewidth]{prospector/color_scale} % 0.7
\caption[The same feature shown as line plot and partial dependence bar.]{
The same feature shown as line plot (top) and partial dependence bar (middle).
Color indicates the predicted risk for the outcome.
The color mapping is shown at the bottom.
}
\label{figs:compress}
\end{figure}