### 2018-04-04 tex/conclusion.tex
a bunch

\chapter{Conclusion \& Future Work}
\label{chap:future_work}

In this thesis we explored how visual analytics can be used in order to leverage the full potential of black-box machine learning.
To that extend we used a variety of techniques to explain the behavior of machine learning models, both globally and locally, and developed a workflow to utilize those explanations to diagnose model \todo{TODO}