### 2018-04-04 tex/conclusion.tex
some changes

\chapter{Conclusion \& Future Work}
\label{chap:future_work}

In this thesis we explored how visual analytics can be used in order to leverage the full potential of black-box machine learning.
To this extend we used a variety of techniques to explain the behavior of machine learning models, both globally and locally, and developed a workflow to utilize those explanations to diagnose models.
We also studied the effect of aggregating local explanations in order to detect biases in the input data to models.
\todo{expand}