### 2018-04-04 vita.tex

%!TEX root = thesis.tex
Josua Walter Hugo Krause was born in Isny im Allg\"au, Germany in January 1988.
He earned his Bachelor's degree in Information Engineering from the University of Konstanz in Germany.
The topic of his Bachelor's thesis was about graphically annotating changes in temporally evolving node link diagrams.
He expanded this topic to using comic book principles in his Master's thesis.
He obtained his Master's degree in Information Engineering also from the University of Konstanz.
In September 2013, he started to pursue his graduate studies towards his Ph.D. under the supervision of Prof. Enrico Bertini.
His research interests were focused on using visualization to support analyzing machine learning.
He published his research on this topic at VAST and CHI, some of the most prestigious conferences in the field of visual analytics and human-computer interaction.
% His works have been cited numerous times.