### 2018-04-06 figs/workflow_intro.tex
typo fixing

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{figs/workflow_large}
\caption[The model diagnostic workflow.]{
The proposed \textit{Model Diagnostic} workflow extends the conventional \textit{Model Building} workflow in machine learning for enabling domain experts to reason about the semantic validity of the decisions made by any model through multiple linked visualizations.
This ultimately helps to improve data acquisition and model generation processes belonging to the original workflow.
}
\label{figs:workflow_intro}
\end{figure}