### 2018-04-06 prospector/related_work.tex

% !TEX root = ../prospector.tex

\section{Related Work}

\subsection{Motivation for Interpretability}
Modern machine learning algorithms are able to create more reliable and precise models
but they are increasingly complex and come with the price of being harder and harder to interpret (Breiman~\cite{breiman2001}).
This inverse relation of understandability versus expressiveness of a model introduces the need
to find ways to improve the interpretability of complex models to overcome this disadvantage.
Lim~\cite{Lim:2012:IUT:2518922} asks questions such as ``\emph{Why} did X happen?",
``\emph{Why} not Y?", ``\emph{What} happens \emph{if} I do Z?", and ``\emph{How} do I make X happen?"
to explain complex mechanisms like machine learning models.
Our system allows users to interactively ask and answer such questions.
Kulesza~\etal~\cite{Kulesza:2015:PED:2678025.2701399} use explanatory debugging by conveying how a model
came to its prediction in order to be able to correct mistakes.
On the other hand, Patel~\etal~\cite{DBLP:conf/ijcai/PatelDFKT11} use multiple classifiers
to better understand input data.
Steeg~and~Galstyan~\cite{NIPS2014_5580,steeg2015corex_theory} use total correlation to build
a hierarchy of features explaining their role in the data.

\subsection{Algorithm Specific Model Visualization}
In the past, research has primarily focused on understanding and interacting with specific machine learning algorithms.
Often the focus is on the internal weights of the trained models.
For Bayesian networks, showing probabilities of the nodes (Becker~\etal~\cite{Becker:2001:VSB:383784.383809}) and how the input is propagated (Correa~\etal~\cite{DBLP:journals/isci/MartinsONHC13})
has been used.
For Support Vector Machines,
projection techniques (Caragea~\etal~\cite{Caragea2001}) and
Nomograms (Jakulin~\etal~\cite{Jakulin:2005:NVS:1081870.1081886})
to see the ``cut" in the instances were utilized.
Visualizing and interpreting the graph of a neural network has also been used by Tzeng and Ma~\cite{Tzeng:2005:OTB}.
Kim~\etal~\cite{kim2014bayesian,kim2015MGM} introduce graphical models that allow for
interpretability of its features.
\cite{kim2015MGM} use a colored matrix of features by category to show distinguishable features computed by their model.
Caruana~\etal~\cite{Caruana:2015:IMH:2783258.2788613} use high-performance generalized additive models (GA$^2$Ms) that allows visual inspection of the influence of its input features on the outcome much like partial dependence.

\subsection{Model Result Visualization}
However, showing only internal, algorithm specific weights is often not enough.
Plate~\etal~\cite{DBLP:journals/neco/PlateBGB00} and Olden~\cite{citeulike:3733836} show how input features influence
the outcome of neural network classifications.
Xu~\etal~\cite{DBLP:journals/corr/XuBKCCSZB15} interprets
the graph of a neural network used for image classification to
retrieve which part of an image was responsible for a specific
classification result.
These techniques aim in the same direction as partial dependence but are limited to only neural networks.
They cannot be used to support the ability to compare different machine learning models across algorithms.

Having access to the internals of a machine learning algorithm also allows direct interaction with the models and to improve them on the fly.
BaobabView (van den Elzen and van Wijk~\cite{van2011baobabview}) offers interactive decision tree construction.
Steed~\etal~\cite{steed2009guided,steed2009tropical}
guides regression model creation by enhancing interaction
with the input data.
EnsembleMatrix (Talbot~\etal~\cite{Talbot:2009:EIV:1518701.1518895})
allows users to combine already computed classification models to
build ensemble models.
Some recent interactive machine learning tools \cite{
amershi15,
Amershi:2011:DEE:2046396.2046416,
Amershi:2012:RIM:2207676.2207680,
Amershi:2011:CHF:1978942.1978966,
Kapoor:2010:IOS:1753326.1753529,
kim2015scalable,
Leung2014710,
Mishra:2015:SAI:2700171.2791022,
export:141330})
are more algorithm agnostic, but depend on general
performance measures like confusion matrices,
area under ROC curve (AUC) measures,
result distribution of instances, and
feature weights according to model independent statistics.

Frank and Hall~\cite{frank2003} use 2D projections of the data to show %classification
results for multiple classification algorithms, as well as Rheingans and desJardins~\cite{885740} with self-organizing maps.

\subsection{Probing Models}
Partial dependence was proposed by Friedman~\cite{friedman2001} for analyzing
gradient boosting models and has since been used for other models as well (\eg, Ehrlinger~\cite{ehrlinger2015} uses partial dependence to analyze the behavior of random forests
in the \emph{R} package \emph{ggRandomForests}).

Cortez and Embrechts~\cite{cortez2011opening,Cortez20131} and Kim~\etal~\cite{kim2014bayesian}
use sensitivity analysis to analyze and compare machine learning models of different algorithms.
Sensitivity analysis is similar to partial dependence except that it uses a few
base vectors (usually the mean, median, or quartiles of all observed values)
instead of computing the probabilities over all instances.
This method is faster than partial dependence but may miss critical details
about the prediction function especially if the function is strongly non-linear.

Goldstein~\etal~\cite{goldstein14} extends the idea of partial dependence by using
Individual Conditional Expectation (ICE) plots which show one line for each row of the input data.
We found, however, that this often clutters the plots too much and makes them harder to interpret.
We experimented further with showing standard deviations and quartiles of the partial dependence line
but discarded this approach since the spread of the partial dependence results is always expected to
be large unless one feature dominates the classification significantly and is able to solely change
the classification even for the furthest instance.

By not restricting ourselves to sampling only the observed input space, our approach on partial dependence
enables a deeper analysis of the machine learning model. Furthermore, accepting the costs of computing
partial dependence over all instances yields proper results even for highly non-linear models,
while also not overwhelming users with too much detail.
This is strengthened even further by our novel approach of using implementation details of the inspected
models to improve the sampling and the representation of the results.